HO unification from object language to meta language

### 2024-03-29 paper/main.tex
update elpi printer for tex
update example eta

$\eta$-reduction is an equivalence relation where a term of the form
$\lambda x.t \appsep x$ can be converted to $t$ any time $x$ does not occur as a
free variable in $t$. We call $\lambda x. t \appsep x$ the $\eta$-expansion of $t$. 
\printAll
  {{\elpiIn{flam x\ fapp [fuva A, x]} \Uo \elpiIn{fcon "f"}}}
  {{\lambda x. A'_x \Ue f}}
  {{\mapping{\elpiIn{A}}{A'}{1}}}
  {{}}
Following the implementation of \elpiIn{comp} given in
\cref{sec:compilation}, the unification problem $\mathcal{P}_1$ given 
above, is compiled into the unification problem $\mathcal{T}_1$. 
We can note that all the terms appearing in $\mathcal{P}_1$ are
unifiable by \Uo, but, despite this, \Ue, is not able to solve $\mathcal{T}_1$:
its left and right hand sides have both different rigid head:
$\lambda x.A'_x$ corresponds to the \Ho term \elpiIn{lam x\ app[con "f", x]} 
and $f$ corresponds to \elpiIn{con"f"}. 
This failure is motivated by the fact that \elpiIn{flam x\ fapp [fuva A, x]}
is a term belonging to \maybeeta.

@@

In order to activate a \linketa $l$, we need to extend the \elpiIn{progress1} predicate
by adding new rules. After passing under all the \elpiIn{abs}
constructors of $l$, there are two cases making a \linketa to progress, 1) \lhs
is instantiated to a rigid term 2) \rhs is no more a \maybeeta or it is a term
which can be reduced to a term with rigid head. 
If \lhs is instantiated to a rigid term $t$, by \cref{inv:subst-nf},
we know that $t$ does not contain any \maybeeta. Let $t'$ the right hand side,
if $t$ is a constant or a function application, then, $t'$, which by
construction has \elpiIn{lam} as head, should be an $\eta$-expansion. We are
therefore allowed to unify $\lambda x. t \appsep x$ (the $\eta$-expanded version
of $t$) with $t'$. Finally, if $t$ is a term with \elpiIn{lam} as head, then
it is not an $\eta$-expansion and therefore, $t$ can be unified with $t'$.

@@

An example

TODO: example for case 1: $\lambda x. \lambda y. F \appsep y \appsep x = f$
TODO: example for case 2: $\lambda x. \lambda y. F \appsep y \appsep x = G, F = \lambda x. \lambda y. a$

>>>>

$\eta$-reduction is an equivalence relation where a term of the form
$\lambda x.t \appsep x$ can be converted to $t$ any time $x$ does not occur as a
free variable in $t$. We call $\lambda x. t \appsep x$ the $\eta$-expansion of $t$. 
\printAll
  {{\elpiIn{flam x\ fapp [fuva X, x]} \Uo \elpiIn{fcon"f"}}}
  {{\lambda x. X'_x \Ue f}}
  {{\mapping{\elpiIn{X}}{X'}{1}}}
  {{}}
Following the implementation of \elpiIn{comp} given in
\cref{sec:compilation}, the unification problem $\mathcal{P}_1$ given 
above, is compiled into the unification problem $\mathcal{T}_1$. 
We can note that all the terms appearing in $\mathcal{P}_1$ are
unifiable by \Uo, but, despite this, \Ue, is not able to solve $\mathcal{T}_1$:
its left and right hand sides have both different rigid head:
$\lambda x.A'_x$ corresponds to the \Ho term \elpiIn{lam x\ app[con "f", x]} 
and $f$ corresponds to \elpiIn{con"f"}. 
This failure is motivated by the fact that \elpiIn{flam x\ fapp [fuva A, x]}
is a term belonging to \maybeeta.

@@

In order to activate a \linketa $l$, we need to extend the \elpiIn{progress1} predicate
by adding new rules. After passing under all the \elpiIn{abs}
constructors of $l$, there are two cases making a \linketa to progress, 1) \lhs
is instantiated to a rigid term 2) \rhs is no more a \maybeeta or it is a term
which can be reduced to a term with rigid head. 
If \lhs is instantiated to a rigid term $t$, by \cref{prop:nf},
we know that $t$ does not contain any \maybeeta. Let $t'$ the right hand side,
if $t$ is a constant or a function application, then, $t'$, which by
construction has \elpiIn{lam} as head, should be an $\eta$-expansion. We are
therefore allowed to unify $\lambda x. t \appsep x$ (the $\eta$-expanded version
of $t$) with $t'$. Finally, if $t$ is a term with \elpiIn{lam} as head, then
it is not an $\eta$-expansion and therefore, $t$ can be unified with $t'$.

@@

\printAll
  {{\elpiIn{flam x\ fapp [fuva X, x] ~\Uo~fcon"f"}}}
  {{A \Ue f}}
  {{\mapping{X}{X'}{1}}}
  {{\linketaM{}{A}{\lambda x.X'_{x}}}}

The example above shows the new compilation of the unification problem given at
the beginning of \cref{sec:eta}. This time, we see that the the left hand side
$t$ of $\mathcal{P}_1$ has been detected as a \maybeeta and replaced with the
fresh variable $A$. Moreover, \linkStore contains the link $l_1$ connecting $A$
with $\lambda x.X'_{x}$, the compiled version of $t$. After the resolution of
$\mathcal{T}_1$, $A$ is assigned to $f$. Therefore \lhs of $l_1$ is a term with
\elpiIn{con} as constructor. This means that \rhs of $l_1$ is a $\eta$-expansion
and therefore we can unify $\lambda x.f x$ with \rhs, which instantiate $X'_x$
to $f \appsep x$. $l_1$ is removed from \linkStore which is now empty,
\elpiIn{progress} terminate, and the decompilation, will instantiate the $X$
variable to $f$ (which is the $\eta$-contracted version of $\lambda x.f \appsep
x$).

A second example, showing the activation of a link when the \rhs is no more a
\maybeeta, is given in \cref{sec:invariant1}, since we need to work with
variables used with different arities.

TODO: example for case 2: $\lambda x. \lambda y. F \appsep y \appsep x = G, F = \lambda x. \lambda y. a$
